How Kids Code and How We Know:\\An Exploratory Study on the Scratch Repository

### 2016-04-22 Paper/paper.tex
updated the dead code smell

		\textit{Projects with:} & & \\
		\hline
		Procedures & 17,979 & 7.70\%\\
		Recursive procedures &  1,052 & 0.45\%\\
		Variables & 73,577 & 31.51\% \\
		Lists & 9,358 & 4.01\% \\
		Conditional statements & 92,959 & 39.81\% \\
		User input blocks& 131,314 & 56.24\% \\
		Loop statements & 180,210 & 77.18\% \\
		\texttt{repeat until <condition>} statements & 31,739 & 13.59\% \\
		\texttt{broadcast} - \texttt{receive} statements & 69,039 & 29.57\%\\
		Cloned scripts across sprites & 59,634 & 25.54\% \\
		Cloned scripts within sprites & 23,671 & 10.14\%\\
		Cloned procedures & 4,945 & 2.12\%\\
		Cloned functionality blocks across sprites & 60,554 & 25.93\%\\
		Exact clones across sprites & 27,574 & 11.81\%\\
		Exact clones within sprites & 2,043 & 0.87\%\\
		Dead code & 56,890	& 24.36\%\\
		Large scripts & 69,521 & 29.77\%\\
		Large sprites & 31,954 & 13.68\%\\		
		\hline
	\end{tabular}
	\caption{Elements and characteristics of the projects in the dataset}
	\label{tbl-characteristics}
\end{table*}

@@

The use of procedures in projects was further investigated through the procedure calls, summarized in Figure \ref{fig:procedurecalls}. A significant number of the defined procedures (13,036 or 5.06\%) are not called in the projects. Most procedures are called exactly once (62.32\% of them) or twice (14.30\%) and from exactly one script (85.92\% of them). Examining the origin of procedure calls, we observed that most of the calls (56.09\%) originate from other procedures, and even 1.06\% originate from the same procedure, making them recursive calls. These recursive procedures are found in 1,052 projects, whose identifiers are made available.\footnotemark[\ref{repo}]

@@

Users can define their own events, using the blocks \texttt{broadcast}, \texttt{broadcast and wait} and \texttt{when I receive}. 
These blocks are used by 29.57\% of the projects. \texttt{broadcast and wait} is rarely used, in only 3.87\% of the projects. The events are not always synchronized: 3,33\% of the \texttt{when I receive} blocks were found to wait for a message that is never being broadcasted, while 4,4\% of the \texttt{broadcast} blocks broadcast a message that is not being received. This lack of syncronizations occurs in 18,669 projects (7.99\% of the total non-empty ones).

%user interativity		
\noindent
\fbox{
	\begin{minipage}{0.46\textwidth}
RQ2: One third of the projects use variables, sometimes without initializing them. 39.81\% of the projects contain conditional statements and 77.18\% contain loops, but conditional loops are rarely used. More than half of the projects are interactive. 29.57\% of the projects use broadcast and receive blocks, but in 7.99\% of the projects they are not synchronized. Most procedures are called once or twice, from a single script which, in more than half of the cases, is another procedure. Recursive procedure calls exist in 1052 (0.45\% of the total) projects.
	\end{minipage}
}

@@

The second smell that we examine is the dead code smell. Already in the previous section, Scratch projects were found to have two types of dead code: defined procedures that are not called (5.06\% of the defined procedures in 2,079 projects) and unmatched broadcast-receive messages (in 7.99\% of the projects). In this section we examine an additional case: scripts that are never invoked due to the lack of a starting \texttt{when <trigger>} block from the Scratch Events or Control category, like in the example in Figure \ref{fig:Unconnected_clones}, or scripts that are comprised of only a \texttt{when <trigger>} block without any functionality. A total of 322,475 scripts like that were found in 56,890 (24.36\% of the total) projects. The majority of these scripts (86.6\%) are scripts missing the starting block. Examining the size of these dead scripts, 72.34\% are composed of a single block. As shown in Figure \ref{fig:deadcode}, however, considerably large dead scripts exist; 2,358 of these scripts originating from 1,553 different projects have more than 30 blocks and up to 2,610.\footnotemark[\ref{repo}]

@@

\begin{figure}
	\begin{center}
		\includegraphics[width=6cm]{fig/Unconnected_clones.png}
		\caption{Cloned blocks that are not connected to an event. Project id: 12237615}
		\label{fig:Unconnected_clones}
	\end{center}
\end{figure}

@@

\noindent
\fbox{
	\begin{minipage}{0.46\textwidth}
RQ3: Code clones are found in 25.54\% of the projects, with 46.55\% of the clones copied three or more times, in the same or across sprites. 24.36\% of the projects contain scripts that are never invoked, and thus exhibit the dead code smell. In some cases these scripts are large. The large script smell is found in 29.77\% of the projects and the large sprite in 13.68\%.
	\end{minipage}}

>>>>

		\textit{Projects with:} & & \\
		\hline
		Procedures & 17,979 & 7.70\%\\
		Recursive procedures &  1,052 & 0.45\%\\
		Variables & 73,577 & 31.51\% \\
		Lists & 9,358 & 4.01\% \\
		Conditional statements & 92,959 & 39.81\% \\
		User input blocks& 131,314 & 56.24\% \\
		Loop statements & 180,210 & 77.18\% \\
		\texttt{repeat until <condition>} statements & 31,739 & 13.59\% \\
		\texttt{broadcast} - \texttt{receive} statements & 69,039 & 29.57\%\\
		Cloned scripts across sprites & 59,634 & 25.54\% \\
		Cloned scripts within sprites & 23,671 & 10.14\%\\
		Cloned procedures & 4,945 & 2.12\%\\
		Cloned functionality blocks across sprites & 60,554 & 25.93\%\\
		Exact clones across sprites & 27,574 & 11.81\%\\
		Exact clones within sprites & 2,043 & 0.87\%\\
		Dead code & 65,760 & 28.16\%\\
		Large scripts & 69,521 & 29.77\%\\
		Large sprites & 31,954 & 13.68\%\\		
		\hline
	\end{tabular}
	\caption{Elements and characteristics of the projects in the dataset}
	\label{tbl-characteristics}
\end{table*}

@@

The use of procedures in projects was further investigated through the procedure calls, summarized in Figure \ref{fig:procedurecalls}. Most procedures are called exactly once (62.32\% of them) or twice (14.30\%) and from exactly one script (85.92\% of them). Examining the origin of procedure calls, we observed that most of the calls (56.09\%) originate from other procedures, and even 1.06\% originate from the same procedure, making them recursive calls. These recursive procedures are found in 1,052 projects, whose identifiers are made available.\footnotemark[\ref{repo}]

@@

Users can define their own events, using the blocks \texttt{broadcast}, \texttt{broadcast and wait} and \texttt{when I receive}.These blocks are used by 29.57\% of the projects. \texttt{broadcast and wait} is rarely used, in only 3.87\% of the projects.

%user interativity		
\noindent
\fbox{
	\begin{minipage}{0.46\textwidth}
RQ2: One third of the projects use variables, sometimes without initializing them. 39.81\% of the projects contain conditional statements and 77.18\% contain loops, but conditional loops are rarely used. More than half of the projects are interactive. 29.57\% of the projects use broadcast and receive blocks. Most procedures are called once or twice, from a single script which, in more than half of the cases, is another procedure. Recursive procedure calls exist in 1052 (0.45\% of the total) projects.
	\end{minipage}
}

@@

The second smell that we examine is the dead code smell. We identify four types of dead code: (1) procedures that are not invoked, (2) unmatched broadcast-receive messages, (3) code that is not invoked and (4) empty event scripts. Investigating the first type, we find that a significant number of the defined procedures (13,036 or 5.06\%) are not called in the projects. This is also shown in Figure \ref{fig:procedurecalls} and it occurs in 2,079 projects. For the second type, we examined the broadcast-receive messages and found that they are not always synchronized: 3,33\% of the \texttt{when I receive} blocks were found to wait for a message that is never being broadcasted, while 4,4\% of the \texttt{broadcast} blocks broadcast a message that is not being received. This lack of syncronizations occurs in 18,669 projects (7.99\% of the total non-empty ones).

The third and fourth cases are detected in scripts that are never invoked due to the lack of a starting \texttt{when <trigger>} block from the Scratch Events or Control category, or scripts that are comprised of only a \texttt{when <trigger>} block without any functionality. A total of 322,475 scripts like that were found in 56,890 (24.36\% of the total) projects. The majority of these scripts (86.6\%) are scripts missing the starting block. Examining the size of these dead scripts, 72.34\% are composed of a single block. As shown in Figure \ref{fig:deadcode}, however, considerably large dead scripts exist; 2,358 of these scripts originating from 1,553 different projects have more than 30 blocks and up to 2,610.\footnotemark[\ref{repo}]

@@

The number of projects exhibiting the dead code smell, considering all four types of dead code, is 65,760 (28.16\% of the total projects).

@@

\noindent
\fbox{
	\begin{minipage}{0.46\textwidth}
RQ3: Code clones are found in 25.54\% of the projects, with 46.55\% of the clones copied three or more times, in the same or across sprites. 28.16\% of the projects contain code that is never invoked, and thus exhibit the dead code smell. In some cases these scripts are large. The large script smell is found in 29.77\% of the projects and the large sprite in 13.68\%.
	\end{minipage}}
